2040 greenhouse gas reduction targets and energy transitions in line with the EU Green Deal

### 2025-11-28 report_paperCharts.Rmd

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
    geom_line(data=h,linewidth=0.7,linetype="dashed") +
    geom_line(data=linearProjection,linewidth=0.8,linetype="dashed") +
    geom_segment(data=linearProjection %>% filter(period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=linearProjection %>% filter(period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_label_repel(data=data.frame(period=c(2042.5),value=mean(linearProjection %>% filter(period %in% c(2040,2045)) %>% pull(value))),
                     label = "linear\nreduction",
                     nudge_x = 0,
                     nudge_y = 2500,
                     segment.square = TRUE,
                     point.padding = 0,
                     hjust = 0.5,
                     direction="y",
                     size = geomText.size) + 
    expand_limits(y=emi1990) +
    scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000)) +
    scale_x_continuous(breaks = seq(1990,2050,10)) +
    labs(title = expression(paste("(a) Total GHG emissions (Mt ", CO[2],"eq/yr)"))) + 
    theme_doubleColumn() +
    theme(axis.title.y=element_blank(),
          plot.title = element_text(margin=margin(b=35)))

@@

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-0.2, hjust=0.5, na.rm=TRUE, size = geomText.size * 0.8, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geomText.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=1.5, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  labs(title = "(b) Emission reductions (%)") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        plot.title = element_text(margin=margin(b=35)))

@@

g[["GHG Emissions per sector"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_area(aes(fill=variable), alpha = 0.8) +
  geom_line(data=totalData, linewidth = 1, color="black") +
  geom_vline(xintercept=2020, color = "black", linetype="dashed", linewidth=1, alpha = 0.5) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars)) +
  scale_x_continuous(breaks=c(2010, 2010, 2020,2030,2040,2050)) + 
  labs(title = "(c) GHG Emissions per sector in the reference scenario (Mt CO2/yr)") + 
  theme_doubleColumn() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.spacing.x = unit(0.2, 'cm'),
        legend.position="right", 
        legend.direction = "vertical", 
        legend.box = "vertical",
        legend.margin=margin(t=30),
        plot.margin = margin(t = 15, unit = "pt"))

@@

g[["2040 Net Emissions - Energy and Industrial Processes - horizontal bar - net only"]] <- plotData %>%
  ggplot(aes(xmid, ymid, fill = ext_Var)) +
  geom_rect(aes(xmin = xmin, xmax = xmax+CCData, ymin = ymin, ymax = ymax,
                alpha = factor(period), color = variable)) +

  geom_errorbar(data=plotData %>% filter(period == 2040), aes(xmin=xmax-value+min+CCData, xmax=xmax-value+max+CCData), width=4, colour="black", alpha=0.9, linewidth=0.8) +

  geom_text(data=plotData %>% filter(period == 1990), aes(y = 2005, label = labels, group = labels, hjust=1, angle = 90), vjust=0.5, hjust=1.1, na.rm=TRUE, size = geomText.size * 0.8, angle = 90, lineheight = 0.8) +

  geom_text(data=plotData %>% filter(period == 2040), aes(x=xmid , y = 2045, label = pct1990), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geomText.size, angle = 90) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) + 
  scale_alpha_manual(values = c(0.2, 0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  scale_y_continuous(breaks=c(2010,2020,2030,2040), labels=c("1990","avrg. 2018-2022","2030","2040"), minor_breaks = NULL, limits=c(1998,2047)) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  #theme(plot.margin = margin(t = 30, r = 0, b = 100, l = 0, unit = "pt")) +
  labs(title = "(d) Sectoral breakdown of residual net CO2 emissions from energy and industrial processes") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        plot.margin = margin(t = 15, r = 0, b = 100, l = 0, unit = "pt"),
        plot.title = element_text(margin=margin(t=0,r=0,b=55,l=0), hjust=1))

@@

ggsave(paste0(outChartsFolder,"/png/Figure 1. Greenhouse gas emissions for scenarios reaching climate neutrality by 2050.png"), g[["Figure 1"]], device="png", width = 12, height = 14, dpi=300, units = "in")
ggsave(paste0(outChartsFolder,"/svg/Figure 1. Greenhouse gas emissions for scenarios reaching climate neutrality by 2050.svg"), g[["Figure 1"]], device="svg", width = 12, height = 14, dpi=100, units = "in")

@@

@@

g[["milestone"]][["Renewables Share"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = Inf, y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.075, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  annotate("text", x = Inf, y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_hline(yintercept=addHist$res %>% filter(type == "RES", period == 2022) %>% pull(value), color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = Inf, y = addHist$res %>% filter(type == "RES", period == 2022) %>% pull(value), label = "2022", angle=90, vjust = -0.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=0, b=20)),
        strip.clip = "off")

@@

  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = -Inf, y = 50, label = "Net-Zero\nIndustry Act", angle=90, vjust = 0.5, hjust = 1, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,300)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=-10, b=20)),
        strip.clip = "off")

@@

g[["milestone"]][["Domestic Hydrogen Production"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = -Inf, y = 10, label = "REPowerEU\ndomestic", angle=90, vjust = 0.5, hjust = 0.8, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,20)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=-15, b=20)),
        strip.clip = "off")
```

@@

g[["Figure 5"]] <- gridExtra::arrangeGrob(
  g[["milestone"]][["Emissions reductions"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Final Energy Demand"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Renewables Share"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Direct Electrification"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Carbon-free electricity share"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Domestic Hydrogen Production"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Electricity Demand"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Wind capacity"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Solar photovoltaics"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Residual Fossil Emissions"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Carbon Capture and Storage"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Novel CDR"]] + theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  ncol = 2
)

@@

@@

### Supplementary figure 1. VRE sensitivity.

<!-- (a) Wind and Solar generation (TWh) -->

```{r supplChart_figure1a, echo=FALSE}

vars <- c(
#  "Solar CSP" = "SE|Electricity|Solar|CSP",
  "Solar PV" = "SE|Electricity|Solar|PV",
  "Wind Onshore" = "SE|Electricity|Wind|Onshore",
  "Wind Offshore" = "SE|Electricity|Wind|Offshore")

color <- c(
  "Solar CSP"     = "#ffcc00",
  "Solar PV"      = "#ffe600",
  "Wind Onshore"  = "#193F7F",
  "Wind Offshore" = "#337fff")

pattern <- c(
  "default"="none",
  "limited VRE"="stripe"
)

plotData <- df %>%
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), scenario %in% c(mainScen,"Nzero_57_bio7p5_limVRE"), variable %in% vars) %>%
  mutate(scenName = ifelse(scenario == mainScen, "default", "limited VRE"),
         value = round(value * EJ2TWh),
         unit = "TWh")

#plotData %>% filter( period == 2040) %>% group_by(scenario) %>% summarise(value = sum(value))
#value2040 <- plotData %>% filter( period == 2040) %>% group_by(scenario) %>% summarise(value = sum(value)) %>% pull(value)
#round((1-(value2040[2]/value2040[1]))*100,2)

barwidth <- 3

g[["Wind and Solar generation"]] <- ggplot() +
  geom_bar_pattern(data = plotData %>% filter(scenario == mainScen), mapping = aes(x=period - (barwidth + 1)/2,y=value, fill=variable, pattern=scenName), stat='identity', position="stack", width = barwidth) +
  geom_bar_pattern(data = plotData %>% filter(scenario == "Nzero_57_bio7p5_limVRE"), mapping = aes(x=period + (barwidth + 1)/2,y=value, fill=variable, pattern=scenName), stat='identity', position="stack", width = barwidth, pattern_key_scale_factor = 0.3, pattern_density=0.3, pattern_spacing = 0.03, pattern_colour = "white", pattern_fill = "white", pattern_alpha=0.4) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(nrow = 1, order = 1, override.aes = list(pattern = "none"))) +
  scale_pattern_manual(values = pattern, breaks=c("default","limited VRE"), guide = guide_legend(nrow = 2, order = 1)) +
  scale_x_continuous(breaks=c(2020,2030,2040,2050), minor_breaks = NULL) +
  labs(title = "(a) Wind and Solar Generation (TWh)") + 
  theme_doubleColumn() +
  theme(axis.title.y = element_blank(),
        legend.margin=margin(t = -0.4, unit='cm'),
        legend.box = "horizontal", 
        legend.box.just = "center")

(b) Limited VRE target reduction (%)

```{r supplChart_figure1b, echo=FALSE}

# error bars for scenarios with and without limVRE

plotData <- rbind(
  df %>%
    filter(!grepl("limVRE", scenario), region == "EU27", period == 2040, tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
    mutate(reduction = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)) ,2),
           scenName = "without\nlim-VRE") %>%
    select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  group_by(period) %>%
  mutate(min = min(reduction),
        max = max(reduction))
  ,
  df %>%
    filter(region == "EU27", period == 2040, tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
    mutate(reduction = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)) ,2),
           scenName = "with\nlim-VRE") %>%
    select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  group_by(period) %>%
  mutate(min = min(reduction),
        max = max(reduction))
)

#plotData %>% filter(scenario %in% c(mainScen,"Nzero_57_bio7p5_limVRE"))

g[["limVRE target reduction"]] <- ggplot(plotData, aes(x = scenName, y = reduction, group=scenName)) +
  geom_errorbar(mapping = aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +
  geom_boxplot(outlier.shape = NA, width=0.3) +
  geom_hline(yintercept=1, color = "black", linewidth=2, alpha = 0.5) +
  scale_y_reverse(breaks=c(1,0.95,0.9,0.85,0.8, 0.75), minor_breaks = NULL, labels = scales::percent) +
  expand_limits(y=c(1,0.75)) +
  labs(title = "(b) 2040 target (%)") +
  theme_doubleColumn() +
  theme(plot.title = element_text(hjust=1))

>>>>

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
    geom_line(data=h,linewidth=0.7,linetype="dashed") +
    geom_line(data=linearProjection,linewidth=0.8,linetype="dashed") +
    geom_segment(data=linearProjection %>% filter(period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=linearProjection %>% filter(period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_label_repel(data=data.frame(period=c(2042.5),value=mean(linearProjection %>% filter(period %in% c(2040,2045)) %>% pull(value))),
                     label = "linear\nreduction",
                     nudge_x = 0,
                     nudge_y = 2500,
                     segment.square = TRUE,
                     point.padding = 0,
                     hjust = 0.5,
                     direction="y",
                     size = geomText.size) + 
    expand_limits(y=emi1990) +
    scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000)) +
    scale_x_continuous(breaks = seq(1990,2050,10)) +
    labs(title = expression(paste("(a) Total GHG emissions (Mt ", CO[2],"eq/yr)"))) + 
    theme_doubleColumn() +
    theme(axis.title.y=element_blank(),
          plot.title = element_text(margin=margin(b=30)))

@@

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-0.2, hjust=0.5, na.rm=TRUE, size = geomText.size * 0.8, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geomText.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=1.5, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  labs(title = "(b) Emission reductions (%)") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        plot.title = element_text(margin=margin(b=30)))

@@

g[["GHG Emissions per sector"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_area(aes(fill=variable), alpha = 0.8) +
  geom_line(data=totalData, linewidth = 1, color="black") +
  geom_vline(xintercept=2020, color = "black", linetype="dashed", linewidth=1, alpha = 0.5) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars)) +
  scale_x_continuous(breaks=c(2010, 2010, 2020,2030,2040,2050)) + 
  labs(title = "(c) GHG Emissions per sector in the reference scenario (Mt CO2/yr)") + 
  theme_doubleColumn() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.spacing.x = unit(0.2, 'cm'),
        legend.position="right", 
        legend.direction = "vertical", 
        legend.box = "vertical",
        legend.margin=margin(t=30),
        plot.margin = margin(t = 15, r=10, unit = "pt"),
        plot.title = element_text(margin=margin(b=20)))

@@

g[["2040 Net Emissions - Energy and Industrial Processes - horizontal bar - net only"]] <- plotData %>%
  ggplot(aes(xmid, ymid, fill = ext_Var)) +
  geom_rect(aes(xmin = xmin, xmax = xmax+CCData, ymin = ymin, ymax = ymax,
                alpha = factor(period), color = variable)) +
  geom_errorbar(data=plotData %>% filter(period == 2040), aes(xmin=xmax-value+min+CCData, xmax=xmax-value+max+CCData), width=4, colour="black", alpha=0.9, linewidth=0.8) +
  geom_text(data=plotData %>% filter(period == 1990), aes(y = 2005, label = labels, group = labels, hjust=1, angle = 90), vjust=0.5, hjust=1.1, na.rm=TRUE, size = geomText.size * 0.8, angle = 90, lineheight = 0.8) +
  geom_text(data=plotData %>% filter(period == 2040), aes(x=xmid , y = 2045, label = pct1990), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geomText.size * 0.8, angle = 90) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) + 
  scale_alpha_manual(values = c(0.2, 0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  scale_y_continuous(breaks=c(2010,2020,2030,2040), labels=c("1990","avrg. 2018-2022","2030","2040"), minor_breaks = NULL, limits=c(1998,2047)) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  #theme(plot.margin = margin(t = 30, r = 0, b = 100, l = 0, unit = "pt")) +
  labs(title = "(d) Sectoral breakdown of residual net CO2 emissions from energy and industrial processes") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        plot.margin = margin(t = 25, r = 0, b = 60, l = 0, unit = "pt"),
        plot.title = element_text(margin=margin(t=0,r=0,b=35,l=0), hjust=1))

@@

ggsave(paste0(outChartsFolder,"/png/Figure 1. Greenhouse gas emissions for scenarios reaching climate neutrality by 2050.png"), g[["Figure 1"]], device="png", width = 12, height = 13, dpi=300, units = "in")
ggsave(paste0(outChartsFolder,"/svg/Figure 1. Greenhouse gas emissions for scenarios reaching climate neutrality by 2050.svg"), g[["Figure 1"]], device="svg", width = 12, height = 13, dpi=100, units = "in")

@@

```{r, echo=FALSE, fig.width=12, fig.height=13}

@@

g[["milestone"]][["Renewables Share"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = Inf, y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.075, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  annotate("text", x = Inf, y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_hline(yintercept=addHist$res %>% filter(type == "RES", period == 2022) %>% pull(value), color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = Inf, y = addHist$res %>% filter(type == "RES", period == 2022) %>% pull(value), label = "2022", angle=90, vjust = -0.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)),
        #strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=0, b=20)),
        strip.clip = "off")

@@

  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = -Inf, y = 50, label = "Net-Zero\nIndustry Act", angle=90, vjust = 0.5, hjust = 1, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,300)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)),
        #strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=-10, b=20)),
        strip.clip = "off")

@@

g[["milestone"]][["Domestic Hydrogen Production"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = -Inf, y = 10, label = "REPowerEU\ndomestic", angle=90, vjust = 0.5, hjust = 0.8, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,20)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)),
        #strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=-15, b=20)),
        strip.clip = "off")
```

@@

g[["Figure 5"]] <- gridExtra::arrangeGrob(
  g[["milestone"]][["Emissions reductions"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Final Energy Demand"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Renewables Share"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Direct Electrification"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Carbon-free electricity share"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Domestic Hydrogen Production"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Electricity Demand"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Wind capacity"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Solar photovoltaics"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Residual Fossil Emissions"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Carbon Capture and Storage"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Novel CDR"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  ncol = 2
)

@@

<!-- ```{r, echo=FALSE, include=FALSE} -->

<!-- g[["tmp"]] <- gridExtra::arrangeGrob( -->
<!--   g[["milestone"]][["Final Energy Demand"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Electricity Demand"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   #g[["milestone"]][["Emissions reductions"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Direct Electrification"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Renewables Share"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Wind capacity"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Solar photovoltaics"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   #g[["milestone"]][["Carbon-free electricity share"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Domestic Hydrogen Production"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   #g[["milestone"]][["Residual Fossil Emissions"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   #g[["milestone"]][["Carbon Capture and Storage"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   g[["milestone"]][["Novel CDR"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)), -->
<!--   ncol = 2 -->
<!-- ) -->

<!-- ggsave(paste0(outChartsFolder,"/svg/tmp.svg"), g[["tmp"]] , device="svg", width = 12, height = 10, dpi=100, units = "in") -->
<!-- ggsave(paste0(outChartsFolder,"/png/tmp.png"), g[["tmp"]] , device="png", width = 12, height = 10, dpi=300, units = "in") -->

<!-- ``` -->

@@

### Supplementary figure 1. EU-27 carbon price

```{r supplChart_figure1a_lineplot, echo=FALSE, include=FALSE}
